Network Configuration Synthesis with Abstract Topologies

### 2016-11-14 paper/abstraction/paper.tex
add early explanation of regex constraints
typo fixes

The second line introduces a prefix \emph{template} variable $\CD{\$GP}$. It says that traffic for each global prefix associated with the variable should follow a path that ends at its corresponding destination router in the $\CD{TG}$ role. The second line does the same for all local prefixes.
%
The final rule matches all other IP prefix destinations and allows traffic to follow a path that leaves the datacenter, ending at some outside location, through neighbors $\CD{N1}$ or $\CD{N2}$ with a preference for leaving through $\CD{N1}$. The $\CD{\Prefer}$~symbol indicates that traffic should satisfy the constraint on the left and resort to the backup (right) only when that is not possible due to network failures.%

@@

The first step for the \sysname compiler is to build the product graph (PG), a data structure that is amenable to joint analysis of the topology and routing policy. As a precursor to this step, we convert the routing policy syntax in \S\ref{sec:overview} to a core language based on regular expressions, shown in Figure~\ref{fig:syntax}.
%
A policy has one or more constraints, each an aggregation or a path constraint. Path constraints have a test on a destination prefix and a list of regular expressions describing network paths. Regular paths are defined over network locations, where each location is either a router inside the operator's network, or an external neighbor connected to that network. Values \In~ and \Out, will match any internal and external location respectively. There are two types of predicates, as shown. A test for a concrete prefix $d.d.d.d/[d..d]$ matches a range of IP prefixes (\EG, 10.0.1.0/[24..32]) where metavariable $d$ represents an integer.

@@

For each regular path constraint $r_i$ from $r_1 ~\Prefer~ .. ~\Prefer~ r_k$, we construct a DFA for the reverse of $r_i$. A DFA for $r_i$ is defined as a tuple ($\Sigma, Q_i, F_i, q_{0_i}, \sigma_i$). The alphabet $\Sigma$ is the set of topology locations, $Q_i$ is the set of states for automaton i, $F_i$ is the set of final states, $q_{0_i}$ is the initial state, and $\sigma_i \colon Q_i \times \Sigma \rightarrow Q_i$ is the state transition function.
%
The PG is a tuple ($G'$, $start$, $P$) where $G' = (V',E')$ is has
vertices $V' \colon V \times Q_1 \times \dots \times Q_j$,
edges $E' \colon V' \times V'$,
a unique starting vertex $start$,
and a preference function $P \colon V' \rightarrow 2^{\set{1, \dots, j}}$, mapping nodes in the PG to a set of path ranks.

@@

The first rule, \infrule{I-out1} applies to a learned invariant of the form $L_m(j,k)$ where the number of outgoing edges from any concrete node in the $m$ role is greater than 0. In the worst case, the largest group of concrete nodes we could hope to reach at the $n$ role would be $e_1$ since all j nodes at the bottom may have outgoing edges to the same concrete nodes at the top. Furthermore, the total number of disjoint paths to the $j$ nodes at the bottom is equal to $j*k$. Since extending the existing disjoint paths with disjoint edges keeps the paths disjoint, and since we can not exceed the current number of disjoint paths to the concrete nodes in role $m$ on the bottom, the largest reachable group for the role $n$ on the top will be $min(j*k,e_1)$. 
%
We conservatively use $1$ for the number of disjoint paths to each node in role $n$, since when $n$ is very large, all reachable nodes in role $m$ might only have a single edge to completely different nodes in $n$.
%However, because we do not know for certain that all $j$ nodes in role $m$ have outgoing edges to the same nodes at the top, we conservatively use $1$ for the number of disjoint paths to each node on top. 
%For example, if $n$ is very large, then each node in role $m$ may have only a single edge to each reachable node in role $n$.

@@

The two rules \infrule{I-mesh1} and \infrule{I-mesh2} handle the case where there is a full mesh between the two roles. This happens when the number of outgoing edges ($e_1$) nodes in role $m$ equals the number of nodes on top ($n$). \infrule{I-mesh1} says that we can find disjoint paths to each node in the top role restricted to the number of disjoint paths we started with. \infrule{I-mesh2} uses the fact that each node in the top role is connected to each node in the bottom role to infer that there can be $j$ disjoint paths to any single node in the top role.

The annotation $mincut(X)$ appearing on an edge in an assertion about the fault tolerance between nodes in two different roles. The rule \infrule{I-mincut} uses such assertions. To each node in role $n$, from a node in role $m$, we can construct at least the minimum of $X$ and $k$ disjoint paths.

@@

The rule \infrule{I-striping} is the most complicated case.
It starts with the invariant $L_m(j,k)$ at role $m$ and can be applied if each edge multiplicity $e_i > 0$ is valid given the constraints. The first inference for role $n$ tries to find the largest reachable group with disjoint paths to each. The idea is similar to the rule \infrule{I-out1}, but is able to use the fact that $e_2 > 0$ to learn more about the structure of the concrete topology. In particular it uses the following inequality, where $g$ represents the size of the group for the role $n$:
%
\[ \begin{array}{c}
  (m-j)*e_1 \geq (n-g)*e_2
\end{array} \]
\noindent
%
The remaining nodes ($m-j$) that are not part of the reachable group in the bottom role, each have $e_1$ outgoing edges and must be able to at least ``fill" the incoming edges for the remaining nodes not in the reachable group at the top ($n-g$) with $e_2$ incoming edges each. Solving the inequality gives the lower bound for $g$ used in Figure~\ref{fig:inference-rules}.

The second part of the rule allows uses a similar idea to reason about the overlap between roles $m$ and $o$ with respect to role $n$. This rule is useful particularly useful for datacenter topologies where routers in one tier of the datacenter often have a particular striping pattern with another tier.

@@

Note that, because each inference rule only applies to directed edges in the PG, the algorithm can not make any inferences about connectivity from the AL role to the S role since there is no directed edge between them. This restriction ensures that the analysis is policy-sensitive.

@@

In this case, the algorithm was able to infer that any node in the spine role has at least 1 disjoint path and any node in the TL node has at least 2 disjoint paths to a router in role TG. Note that, in this case, the analysis is precise. There exists a concrete network, namely the datacenter from Figure~\ref{fig:example2}, where a single failure can disconnect a global tor from a spine router due to the valleyfree constraint. In this case in particular, the bound on the fault tolerance between a tor and spine router can be used to determine that an aggregation-induced black hole may arise at the border of the datacenter after a single failures in some topologies.

>>>>

The second line introduces a prefix \emph{template} variable $\CD{\$GP}$. It says that traffic for each global prefix associated with the variable should follow a path that ends at its corresponding destination router in the $\CD{TG}$ role. As we will see in \S\ref{sec:language}, constraints like $\CD{\End(TG)}$ are just syntactic sugar for regular expressions describing network paths. The second line has a similar policy for local prefixes.
%
The final rule matches all other IP prefix destinations and allows traffic to follow a path that leaves the datacenter, ending at some outside location, through neighbors $\CD{N1}$ or $\CD{N2}$ with a preference for leaving through $\CD{N1}$. The $\CD{\Prefer}$~symbol indicates that traffic should satisfy the constraint on the left and resort to the backup (right) only when that is not possible due to network failures.%

@@

The first step for the \sysname compiler is to build the product graph (PG), a data structure that is amenable to joint analysis of the topology and routing policy. As a precursor to this step, we convert the routing policy syntax in \S\ref{sec:overview} to a core language based on regular expressions, shown in Figure~\ref{fig:syntax}.
%
A policy has one or more constraints, each an aggregation or a path constraint. Path constraints have a test on a destination prefix and a list of regular expressions describing network paths. Regular paths are defined over network locations, where each location is either a router inside the operator's network, or an external neighbor connected to that network. Values \In\ and \Out\ will match any internal and external location respectively. There are two types of predicates, as shown. A test for a concrete prefix $d.d.d.d/[d..d]$ matches a range of IP prefixes (\EG, 10.0.1.0/[24..32]) where metavariable $d$ represents an integer.

@@

For each regular path constraint $r_i$ from $r_1 ~\Prefer~ .. ~\Prefer~ r_k$, we construct a DFA for the reverse of $r_i$. A DFA for $r_i$ is defined as a tuple ($\Sigma, Q_i, F_i, q_{0_i}, \sigma_i$). The alphabet $\Sigma$ is the set of topology locations, $Q_i$ is the set of states for automaton i, $F_i$ is the set of final states, $q_{0_i}$ is the initial state, and $\sigma_i \colon Q_i \times \Sigma \rightarrow Q_i$ is the state transition function.
%
The PG is a tuple ($G'$, $start$, $P$) where $G' = (V',E')$ has
vertices $V' \colon V \times Q_1 \times \dots \times Q_j$,
edges $E' \colon V' \times V'$,
a unique starting vertex $start$,
and a preference function $P \colon V' \rightarrow 2^{\set{1, \dots, j}}$, mapping nodes in the PG to a set of path ranks.

@@

The first rule, \infrule{I-out1} applies to a learned invariant of the form $L_m(j,k)$ where the number of outgoing edges from any concrete node in the $m$ role is greater than 0. In the worst case, the largest group of concrete nodes we could hope to reach at the $n$ role would be $e_1$ since all j nodes at the bottom may have outgoing edges to the same concrete nodes at the top. Furthermore, the total number of disjoint paths to the $j$ nodes at the bottom is equal to $j*k$. Since extending the existing disjoint paths with disjoint edges keeps the paths disjoint, and since we cannot exceed the current number of disjoint paths to the concrete nodes in role $m$ on the bottom, the largest reachable group for the role $n$ on the top will be $min(j*k,e_1)$. 
%
We conservatively use $1$ for the number of disjoint paths to each node in role $n$, since when $n$ is very large, all reachable nodes in role $m$ might only have a single edge to completely different nodes in $n$.
%However, because we do not know for certain that all $j$ nodes in role $m$ have outgoing edges to the same nodes at the top, we conservatively use $1$ for the number of disjoint paths to each node on top. 
%For example, if $n$ is very large, then each node in role $m$ may have only a single edge to each reachable node in role $n$.

@@

The two rules \infrule{I-mesh1} and \infrule{I-mesh2} handle the case where there is a full mesh between the two roles. This happens when the number of outgoing edges ($e_1$) from nodes in role $m$ equals the number of nodes on top ($n$). \infrule{I-mesh1} says that we can find disjoint paths to each node in the top role restricted to the number of disjoint paths we started with. \infrule{I-mesh2} uses the fact that each node in the top role is connected to each node in the bottom role to infer that there can be $j$ disjoint paths to any single node in the top role.

The annotation $mincut(X)$ appearing on an edge is an assertion about the fault tolerance between nodes in two different roles. The rule \infrule{I-mincut} uses such assertions. To each node in role $n$, from a node in role $m$, we can construct at least the minimum of $X$ and $k$ disjoint paths.

@@

The rule \infrule{I-striping} is the most complicated case.
It starts with the invariant $L_m(j,k)$ at role $m$ and can be applied if each edge multiplicity $e_i > 0$ is valid given the constraints. The first inference for role $n$ tries to find the largest reachable group with disjoint paths to each. The idea is similar to the rule \infrule{I-out1}, but is able to use the fact that $e_2 > 0$ to learn more about the structure of the concrete topology. In particular it uses the following inequality, where $g$ represents the size of the group for the role $n$:
%
\[ \begin{array}{c}
  (m-j)*e_1 \geq (n-g)*e_2
\end{array} \]
\noindent
%
The remaining nodes ($m-j$) that are not part of the reachable group in the bottom role, each have $e_1$ outgoing edges and must be able to at least ``fill" the incoming edges for the remaining nodes not in the reachable group at the top ($n-g$), which each have $e_2$ incoming edges. Solving the inequality gives the lower bound for $g$ used in Figure~\ref{fig:inference-rules}.

The second part of the rule uses a similar idea to reason about the overlap between roles $m$ and $o$ with respect to role $n$. This rule is useful particularly useful for datacenter topologies where routers in one tier of the datacenter often have a particular striping pattern with another tier.

@@

Note that, because each inference rule only applies to directed edges in the PG, the algorithm cannot make any inferences about connectivity from the AL role to the S role since there is no directed edge from AL to S. This restriction ensures that the analysis is policy-sensitive.

@@

In this case, the algorithm was able to infer that any node in the spine role has at least 1 disjoint path and any node in the TL node has at least 2 disjoint paths to a router in role TG. Note that, in this case, the analysis is precise. There exists a concrete network, namely the datacenter from Figure~\ref{fig:example2}, where a single failure can disconnect a global ToR from a spine router due to the valley-free constraint. In this case in particular, the bound on the fault tolerance between a ToR and spine router can be used to determine that an aggregation-induced black hole may arise at the border of the datacenter after a single failure in some topologies.
